The SMT Competition 2015--2018

### 2019-08-16 2018/report/Reviews_67398_Revised_Response.txt

> *** Review 1 ***
>
>  The paper overall improved since the first submission, but there are still issues to be fixed, some of them already reported in the review of the first version.
>
>  * Organization
>
> - information about ranking should all stay in the related section, while
> there are related information, e.g. in the first part of Section 7 - Results.

TODO

> - outline of the content of Section 8 describes only part of its content (the first three sub-sections).

TODO

> * Quality
>
> - The paper still contains a number of issues, minor or not, detailed at the end of this review.
>
> -  In Section 2 the fact that "appointment should happen ... " is of relative interest and could be omitted. The same holds for the day in which the events took place.

TODO

> * Rules
>
> - basically incorrect results are "ignored" by omitting the benchmark if the status is not known - this comment still remains. The authors updated the text. I see their points and analysis, which I appreciated, through I think that this is an issue that should be dealt with more care. Correctness is very important for solvers when checking in particular real benchmarks (HW- and SW-verification, etc).

TODO

> * Significance of the results
>
>  - the paper and its analysis now contain many numbers and data, which is very appropriate for e competition report in my opinion. But, basically, no explanation is in place. What I expect to get out of a report, other than commenting the numbers, is their explanation, e.g.: what are the techniques implemented in solvers that overall performed better? Rr better, what techniques implemented in solvers led to an efficient solvers on a particular benchmark? why such techniques led to positive results on benchmarks with (a) given structure/characteristics?

TODO

>  - last paragraph of Section 7.1 that gives motivations for MathTSAT and Z3 to be industrial-strenght tool, and states Z3 training, are either not clear or somehow not convincing as motivations. The paragraph should be clarified, or can be omitted.

TODO

>  - Table 16 should not consider, in my opinion, divisions in which there was only one sound solvers, otherwise the concept of "solved uniquely" is trivially true.

TODO

>  * Relation to other competitions
>
>  - in the conclusions section authors mention the changes that will be implemented in 2019: it would be interesting to state which of the changes mentioned are already in place in other related solver's competitions (e.g., SAT, ASP, QBF, SV ...) .

TODO

>  * Minor points
>
>  Introduction
>
>  - "Hence there is" ->  - "Hence there is" ->

WONTFIX (comment is unclear)

>  - there is reference broken at pg. 2

DONE (http:// instead of https:// for www.satcompetition.org)

>  Section 2
>
>  - does the event was co-located with any conference in particular in 2018?

DONE (affiliated with IJCAR in 2018)

>  - "permitted" -> "allowed"

DONE

>  Section 3
>
>  - "increased by" -> "increased of" ?

WONTFIX

>  Section 4
>
>  - "discontinued partly" -> "partly discontinued" ?

DONE (added parentheses for clarity)

>  - "lacking tool support" -> "lacking of tool support" (This recurs later.)

DONE (lacking -> inadequate)

>  - "whose precise" ->  ", whose precise"

DONE

>  - "DT" looks not to be introduced

TODO

>  - the concept of "benchmarks families" is used at the beginning of pg. 7 but it has not been introduced

TODO

>  - "and in 2018, there were" ->   "while in 2018, there were"

DONE

> - e \in \{0,1 \} is defined, but than it is said to be the number of returned statuses...are they aligned? Please clarify

TODO

> Section 6
>
> - "executing" -> "performing"

DONE (executing -> exercising)

> - there is an huge space at pg. 14

TODO

> - Table 9: there are two team members for solver STP that miss affiliation

TODO

> - "Table 11 summarizes these figures". It would be probably useful to explicitly mention the figure numbers

DONE (figures -> numbers)

> Section 7
>
> - "are also taking" -> "were taken"

DONE

> - referring in the text to colors may not be a good choice for b/w printing

TODO

> - In Sec. 7.1 "assume was trained" -> "assume that it was trained"

DONE

> Section 8
>
> - "the fact that in 2015," ->  "the fact that, in 2015,"

DONE

>  - "degression of Z3" ?

DONE

>  - "rank highly" -> "rank in high position" ? (This recurs later.)

DONE

>  - "has relatively few" -> "has few"

DONE

>  - "quotient" -> "ratio" ?

TODO

>  - "a benchmarks" ->  "a benchmark"

DONE

>  - "problem" and "benchmark" are used for the same concept here. Maybe they can be unified

TODO

>  - in Table 17, does 1.0 means they are sequential? In case these numbers could be omitted

WONTFIX (we do not know for sure if those solvers are sequential; and
with these numbers present, the table also indicates whether a solver
was submitted in a given year)

>  - "As in the previous, this" -> "As in the previous analysis, this"

DONE

>  - "Unsurprisingly the" ->   "Unsurprisingly, the"

DONE

>  - "and not Vampire." -> - "and not by Vampire."

DONE

>  Section 9
>
> - "The last FLoC Olympiad was a period of ..." -> are you referring to "the time between the two FLoC Olympiads" ?

WONTFIX (yes; and this is correct usage of the word Olympiad)

>  References
>
>  They could be made more uniform, e.g.
>
>  - [7] vs. [12], where the same journal is referred with contracted form or full name

DONE

>  - Isn't [13] a CADE paper? In case, the way the conference name is mentioned is very different to [14], which is a CADE paper as well

DONE

>  - [38] contains the only acronym for journal name

DONE

> ============================================
>
> *** Review 2 ***
>
> ummary: This is a much improved revision of this paper. I'm very happy to see
> the additional sections and paragraphs, especially Sec 8! I think this paper is
> ready for publication, modulo a few minor (essentially typesetting) issues (see
> below).
>
> Details:
>
> 4.1 Scoring: A good addition. I think it would be great if the changes to the
> scoring system over the years would be analyzed, to see whether they did indeed
> have the intended effect. This has been started in 8.6, but a more
> comprehensive analysis would be great (possibly as a separate submission here
> or elsewhere).

TODO

> (28), Tbl 7: Thanks for the revision of the table, but I'm not sure listing
> more team members to "the best of our knowledge" is a good idea. If I were to
> do that, I would be bound to miss someone... Maybe just list the main
> institution? Or perhaps ask the submitter who they would want to be listed? I
> think there should be enough time to do that.

TODO

> General: Gray and grey are the worst colors to use. On my (Xerox color-laser)
> printer, all of the gray text in tables and figures are almost or completely
> unreadable. For instance, on white (bleached) paper there is not even a faint
> trace of the gray vertical line in Figures 3/4; I had to switch to the screen
> to see what this refers to.

TODO

> Fig 5/6: These could be significantly smaller :)

TODO

> Fig 3/4: Starting the X-axis at an arbitrary point is always a bit fishy. This
> way, any tiny difference can be overemphasized to achieve any level of
> impressiveness that the author choses. For instance, 144k vs 136k instances
> (casually read off the plot) is a 3.5% increase, but this is an irrelevant
> detail as explained in the text. All four series are also of exactly the same
> shape. I think a cactus plot is simply not the right tool for this job.

TODO

> ============================================
>
> *** Review 3 ***
>
>
> This is a revised version of a paper describing
> the years 2015-2018 of the SMT competition.
> The important changes requested last time were addressed,
> and new content was added in Section 8,
> which discusses interesting questions
> and adds substantial value to the paper.
>
> There are only a few minor improvements that should be made before publication:
>
> In the definition of the raw benchmark score (Sect. 4.1),
> it is not clear why CPU time is bounded at 4T
> (the fact that only 4 cores are available only comes later).

TODO

> In the remainder of Sect. 4.1, the term "benchmark family" is used,
> but this is never defined.
> This could be added to the definitions part in the beginning of Sect. 4.

TODO

> In the definition of "Competition-Wide Score (Main Track)",
> it says "since the weight already takes".
> This seems like the weighted division score is used as input for this computation,
> but another sentence of this definition explicitly uses
> the solver's raw score for this computation.
> So what weight is referred to here?

TODO

> In tables 3-5, please keep the regular numbers right-aligned.
> In LaTeX, the easiest way to do this is probably to use separate columns
> for the regular numbers and the number in brackets,
> and potentially shrink the space between each pair of columns.

TODO

> It is said that "StarExec ensures the reproducibility of the competition".
> If we use the terminology of ACM
> (https://www.acm.org/publications/policies/artifact-review-badging),
> the better term would be "replicability" instead of "reproducibility",
> because StarExec allows a different team to replicate the results
> with the same experimental setup.

DONE

> In the next sentence, commas should be placed before and after "e.g.".

DONE

> For tables 7-9, a smaller font size could be used
> (like in Table 12).

TODO

> The first paragraph of Sect. 8 mentions only three aspects,
> but more aspects are analyzed.

TODO

> Figure 5 seems unnecessary, because the difference is not visible anyway.
> Just listing the numbers would be sufficient.

TODO

> Especially in Sect. 8, there are many larger numbers.
> For readability, it would be better if all numbers larger than 9999
> would use a thin space as thousand separator,
> e.g., $1\,000\,000$ or using siunitx which can do this automatically
> with \num{1000000}.
> (Please do not use other characters as thousand separator
> because this can be confusing for international readers.)

TODO

> In Sect. 8.4, it would be better to refer to bounded model checking
> instead of just model checking.
> Other model checking algorithms use many SMT queries
> where the discussed property does not hold.

DONE

> On the middle of page 2, there is an unresolved reference ([?]).

DONE

> Footnote 1 is on the wrong page.

TODO

### 2019-08-19 2018/report/SMT-COMP.tex
Focussing -> Focusing
model checking -> bounded model checking
Participants tables (minor)
Use siunitx's \num command

\usepackage{jsat}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{booktabs}
%\usepackage[pdf]{pstricks}
%\usepackage{pst-tree}
\usepackage{xspace}
\usepackage{enumitem}

@@

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec.  Users of SMT solvers are encouraged to submit new and
interesting benchmarks to the library, which has grown from 1,352
benchmarks in~2005~\cite{BdMS05} to 347,011 benchmarks in 2018.

@@

The 2018 release of the SMT Library contained 336,844 non-incremental
benchmarks in 51 logics, as well as~10,167 incremental benchmarks
(with a total of nearly 32 million {\tt check-sat} commands) in 26
logics.  Table~\ref{table:smtlib} presents further data on the size of
the SMT Library since~2015.  While the library usually grows from one
year to the other because of the inclusion of new benchmarks and
logics, continuous curation efforts may also cause benchmarks to be
removed or reclassified on occasion, thereby causing individual logics
or even the entire library to shrink.  From~2015 to~2018, the total
number of benchmarks has increased by~68\,\%.  The largest benchmark
in the 2018 release had a size of 1,041~MB; the total size of the
release was 91~GB.

@@

\paragraph{Eligible Benchmarks.}
Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, 11,966 benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division), whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
1,104 benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of inadequate tool support.  In~2018, 3,603~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

\paragraph{Benchmark Weights (\maintrack and \ucoretrack 2016-2018).} The \emph{weight} of a
benchmark in the main track is computed as $\alpha = (1 + \ln F)/F$,
where~$F\geq 1$ is the size of the benchmark's family.  We then define its
\emph{normalized weight} as $\alpha' = \alpha / (\sum_{i\in F} \alpha_i)$,
where the sum is over all benchmarks in the division.  The log scaling seems a
reasonable compromise between linearly combining numbers of benchmarks, which
would overemphasize large families, and simply averaging the fraction of
benchmarks solved, which would overemphasize small families.

\paragraph{Weighted Division Scores (\maintrack and \ucoretrack 2016-2018).}
The \emph{(weighted) score} of a solver in a division of the \maintrack and
\ucoretrack with $M$ benchmarks is then defined as the weighted sum of all raw
scores $$\sum_{b\in M} \alpha'_b \cdot \langle e_b \cdot N, n_b \cdot N,
w_b, c_b\rangle$$ where the sum is over all benchmarks in the division to
assess parallel performance, and the weighted sum of all sequential scores to
assess sequential performance.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver/benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~1,028,615 job pairs in~2015 to~1,776,062 job
pairs in~2018.  (This is the number of unique solver/benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~339,714 job pairs~\cite{CDW14}.

@@

\begin{figure}
  \centering
  \scalebox{.45}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2018.pdf}}
  \caption{VBS on common benchmarks 2015-2018 of all logics in the \maintrack
  with a time limit of 1200s. Number of common benchmarks is indicated with a
  gray vertical line.}
  \label{fig:progress-all}
\end{figure}

@@

\begin{figure}
  \centering
  \scalebox{.45}{\includegraphics{%
  plot/pdf/bar_plot-Virtual_Best_Solver_(VBS)_Unsat_Core_Track_2016-2018.pdf}}
  \vspace{-5ex}
  \caption{VBS in terms of number of unsat core reductions
  within a time limit of 2400s
  on common benchmarks 2016-2018 of all logics in the \ucoretrack.}
  \label{fig:progress-all-uc}
\end{figure}

@@

For the \emph{\maintrack}, we determine the VBS as the best wall-clock
performance per (correctly solved) instance with the time limit of 1200s used
in the competitions 2017 and 2018. Figure~\ref{fig:progress-all} visualizes VBS
performance on common benchmarks from 2015-2018 as cactus plot.
Figure~\ref{fig:progress} shows VBS performance on common benchmarks over pairs
of consecutive years from 2014 to 2018.  Both plots show the wall-clock solving
time per instance over all instances, sorted by solving time.
%
Between 2014 (67426 benchmarks) and 2015 (154238 benchmarks),
the number of common benchmarks is 67070;
between 2015 and 2016 (154424 benchmarks) it is 154238;
between 2016 and 2017 (238758 benchmarks) it is 145236;
and between 2017 and 2018 (333241 benchmarks) it is 226429.
The number of solved instances on these sets of common benchmarks
increased from 2014 to 2015 by 0.1\%, from 2016 to 2017 by 3.9\%,
and from 2017 to 2018 by 0.3\%.
Interestingly, it decreased from 2015 to 2016 by 3.5\%.
This is mainly due to the fact that, in 2015, more than 5600 benchmarks of
division QF\_FP where solved by one of the two participating configurations of
Z3 that no one else could solve, and in 2016 Z3 did not participate in this
division.

For the \emph{\apptrack}, we determine the VBS as the best performance
in terms of number of correct \texttt{check-sat} answers within the time limit
of 2400s (the time limit of the \apptrack in the competitions
2015-2018).
The total number of eligible \texttt{check-sat} queries (i.e., all queries with
status sat or unsat up to the first query with status unknown) on
common benchmarks from 2015-2018 is 30370517.
Figure~\ref{fig:progress-all-app} visualizes VBS performance on these
benchmarks as a bar plot.
In 2015, already 99.95\% of the eligible queries were answered correctly,
which naturally does not leave a lot of room for improvement.
In 2016, performance decreased by 185625 correct answers (0.61\%),
which can be entirely attributed to a performance regression
of Z3 on 2 benchmarks in the QF\_NIA division.
%(on which in 2016, CVC4 then
%performed best, and from 2017 onwards even as well as Z3 in 2015).

For the \emph{\ucoretrack}, we determine the VBS as the best performance in
terms of \emph{reduction}, i.e., the difference in the number of assertions
of the input formula and the unsat core, within the time limit
of 2400s (the time limit of the \ucoretrack in the competitions
2016-2018).
Figure~\ref{fig:progress-all-app} visualizes VBS performance on the common
benchmarks (91455 total) from 2016-2018 as a bar plot.
For 32377 of these benchmarks, extracting an unsat core is trivial since
they contain only one single \texttt{assert} command.
They are thus also not considered in the bar plot, since their maximum
possible reduction is trivially 0.

@@

\subsection{Focussing on Satisfiability or Unsatisfiability}

In many cases, an SMT application produces either mainly satisfiable or mainly
unsatisfiable queries to the SMT solver.  A typical example is model checking,
where a query is only satisfiable if a property of the model does not hold.  As
a consequence, some solvers may implement specialized techniques and
optimizations for the more frequent (the expected) case.
%
Therefore, it is interesting to ask whether the winners of divisions of a track
might have changed if benchmarks in that division had been restricted to one of
these results.  Tables~\ref{tab:results:sat}-\ref{tab:results:unsat} show the
changes in division winners if only satisfiable or unsatisfiable results are
given. This table is in reference to Table~\ref{tab:winners-main} and only
lists a solver if the result has changed. A non-competing solver is listed by
itself if it was not the overall best performing solver of that division but
now outperforms the winning competitive solver, and the winning competing
solver did not change.
It is interesting to note that we see some new solvers winning divisions that did not win any division in the main results. For example, Alt-Ergo solves the most satisfiable problems in AUFLIA and ALIA in 2018, and SMTInterpol performs very well on unsatisfiable problems in QF\_LIA.

@@

Whilst producing these results we noticed that the description of what constitutes a family had been incorrectly interpreted by the scoring scripts in 2016-2018. After fixing this misinterpretation we note that only a single division winner changes---in 2017 AUFNIRA should have been won by CVC4 and not by Vampire.

@@

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~77,352
in~2014~\cite{CDW14} to 342,498 in 2018.  This is partly because new
benchmarks were added to the SMT Library, but also because benchmarks
with unknown status are now eligible for the competition, and because
the competition now uses all eligible benchmarks (rather than a
randomly selected subset).

Since the re-introduction of the \ucoretrack in~2016, the
competition again has three separate tracks.  This, together with new
logics in the SMT Library, has brought the total number of competition
divisions to 115 in~2018, up from just 40 divisions
in~2014~\cite{CDW14}.  The number of job pairs increased from 339,714
in~2014 to~1,776,062~in 2018, and the total wall-clock time in~2018
exceeded 7.6~years.  By these measures, the SMT Competition is one of
the largest competitions in automated reasoning.

>>>>

\usepackage{jsat}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{booktabs}
%\usepackage[pdf]{pstricks}
%\usepackage{pst-tree}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{siunitx}

@@

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec.  Users of SMT solvers are encouraged to submit new and
interesting benchmarks to the library, which has grown from \num{1352}
benchmarks in~2005~\cite{BdMS05} to \num{347011} benchmarks in 2018.

@@

The 2018 release of the SMT Library contained \num{336844} non-incremental
benchmarks in 51 logics, as well as~\num{10167} incremental benchmarks
(with a total of nearly 32 million {\tt check-sat} commands) in 26
logics.  Table~\ref{table:smtlib} presents further data on the size of
the SMT Library since~2015.  While the library usually grows from one
year to the other because of the inclusion of new benchmarks and
logics, continuous curation efforts may also cause benchmarks to be
removed or reclassified on occasion, thereby causing individual logics
or even the entire library to shrink.  From~2015 to~2018, the total
number of benchmarks has increased by~68\,\%.  The largest benchmark
in the 2018 release had a size of \num{1041}~MB; the total size of the
release was 91~GB.

@@

\paragraph{Eligible Benchmarks.}
Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, \num{11966} benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division), whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
\num{1104} benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of inadequate tool support.  In~2018, \num{3603}~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

\paragraph{Benchmark Weights (\maintrack and \ucoretrack 2016--2018).} The \emph{weight} of a
benchmark in the main track is computed as $\alpha = (1 + \ln F)/F$,
where~$F\geq 1$ is the size of the benchmark's family.  We then define its
\emph{normalized weight} as $\alpha' = \alpha / (\sum_{i\in F} \alpha_i)$,
where the sum is over all benchmarks in the division.  The log scaling seems a
reasonable compromise between linearly combining numbers of benchmarks, which
would overemphasize large families, and simply averaging the fraction of
benchmarks solved, which would overemphasize small families.

\paragraph{Weighted Division Scores (\maintrack and \ucoretrack 2016--2018).}
The \emph{(weighted) score} of a solver in a division of the \maintrack and
\ucoretrack with $M$ benchmarks is then defined as the weighted sum of all raw
scores $$\sum_{b\in M} \alpha'_b \cdot \langle e_b \cdot N, n_b \cdot N,
w_b, c_b\rangle$$ where the sum is over all benchmarks in the division to
assess parallel performance, and the weighted sum of all sequential scores to
assess sequential performance.

@@

Jobs consist of individual \emph{job pairs}.  Each job pair
corresponds to a specific solver/benchmark combination.
Figure~\ref{fig:job-pairs} shows the total number of job pairs for the
SMT Competition from~2015 to~2018.  Over this period, the number has
grown by~72\,\%, from~\num{1028615} job pairs in~2015 to~\num{1776062} job
pairs in~2018.  (This is the number of unique solver/benchmark
combinations that were part of the competition.  The number of job
pairs that were actually run is even higher: test jobs and job pairs
that had to be re-run, e.g., because of issues with the StarExec
infrastructure, are not included.)  For comparison, the~2014
competition only ran~\num{339714} job pairs~\cite{CDW14}.

@@

\begin{figure}
  \centering
  \scalebox{.45}{\includegraphics{%
        plot/pdf/cactus_plot-Virtual_Best_Solver_(VBS)_Main_Track_2015-2018.pdf}}
  \caption{VBS on common benchmarks 2015--2018 of all logics in the \maintrack
  with a time limit of 1200s. Number of common benchmarks is indicated with a
  gray vertical line.}
  \label{fig:progress-all}
\end{figure}

@@

\begin{figure}
  \centering
  \scalebox{.45}{\includegraphics{%
  plot/pdf/bar_plot-Virtual_Best_Solver_(VBS)_Unsat_Core_Track_2016-2018.pdf}}
  \vspace{-5ex}
  \caption{VBS in terms of number of unsat core reductions
  within a time limit of 2400s
  on common benchmarks 2016--2018 of all logics in the \ucoretrack.}
  \label{fig:progress-all-uc}
\end{figure}

@@

For the \emph{\maintrack}, we determine the VBS as the best wall-clock
performance per (correctly solved) instance with the time limit of 1200s used
in the competitions 2017 and 2018. Figure~\ref{fig:progress-all} visualizes VBS
performance on common benchmarks from 2015-2018 as cactus plot.
Figure~\ref{fig:progress} shows VBS performance on common benchmarks over pairs
of consecutive years from 2014 to 2018.  Both plots show the wall-clock solving
time per instance over all instances, sorted by solving time.
%
Between 2014 (\num{67426} benchmarks) and 2015 (\num{154238} benchmarks),
the number of common benchmarks is \num{67070};
between 2015 and 2016 (\num{154424} benchmarks) it is \num{154238};
between 2016 and 2017 (\num{238758} benchmarks) it is \num{145236};
and between 2017 and 2018 (\num{333241} benchmarks) it is \num{226429}.
The number of solved instances on these sets of common benchmarks
increased from 2014 to 2015 by 0.1\%, from 2016 to 2017 by 3.9\%,
and from 2017 to 2018 by 0.3\%.
Interestingly, it decreased from 2015 to 2016 by 3.5\%.
This is mainly due to the fact that, in 2015, more than \num{5600} benchmarks of
division QF\_FP where solved by one of the two participating configurations of
Z3 that no one else could solve, and in~2016 Z3 did not participate in this
division.

For the \emph{\apptrack}, we determine the VBS as the best performance
in terms of number of correct \texttt{check-sat} answers within the time limit
of 2400s (the time limit of the \apptrack in the competitions
2015--2018).
The total number of eligible \texttt{check-sat} queries (i.e., all queries with
status sat or unsat up to the first query with status unknown) on
common benchmarks from 2015--2018 is \num{30370517}.
Figure~\ref{fig:progress-all-app} visualizes VBS performance on these
benchmarks as a bar plot.
In 2015, already 99.95\% of the eligible queries were answered correctly,
which naturally does not leave a lot of room for improvement.
In 2016, performance decreased by \num{185625} correct answers (0.61\%),
which can be entirely attributed to a performance regression
of Z3 on 2 benchmarks in the QF\_NIA division.
%(on which in 2016, CVC4 then
%performed best, and from 2017 onwards even as well as Z3 in 2015).

For the \emph{\ucoretrack}, we determine the VBS as the best performance in
terms of \emph{reduction}, i.e., the difference in the number of assertions
of the input formula and the unsat core, within the time limit
of 2400s (the time limit of the \ucoretrack in the competitions
2016--2018).
Figure~\ref{fig:progress-all-app} visualizes VBS performance on the common
benchmarks (\num{91455} total) from 2016--2018 as a bar plot.
For \num{32377} of these benchmarks, extracting an unsat core is trivial since
they contain only one single \texttt{assert} command.
They are thus also not considered in the bar plot, since their maximum
possible reduction is trivially 0.

@@

\subsection{Focusing on Satisfiability or Unsatisfiability}

In many cases, an SMT application produces either mainly satisfiable or mainly
unsatisfiable queries to the SMT solver.  A typical example is bounded model checking,
where a query is only satisfiable if a property of the model does not hold.  As
a consequence, some solvers may implement specialized techniques and
optimizations for the more frequent (the expected) case.
%
Therefore, it is interesting to ask whether the winners of divisions of a track
might have changed if benchmarks in that division had been restricted to one of
these results.  Tables~\ref{tab:results:sat}-\ref{tab:results:unsat} show the
changes in division winners if only satisfiable or unsatisfiable results are
given. This table is in reference to Table~\ref{tab:winners-main} and only
lists a solver if the result has changed. A non-competing solver is listed by
itself if it was not the overall best performing solver of that division but
now outperforms the winning competitive solver, and the winning competing
solver did not change.
It is interesting to note that we see some new solvers winning divisions that did not win any division in the main results. For example, Alt-Ergo solves the most satisfiable problems in AUFLIA and ALIA in 2018, and SMTInterpol performs very well on unsatisfiable problems in QF\_LIA.

@@

Whilst producing these results we noticed that the description of what constitutes a family had been incorrectly interpreted by the scoring scripts in 2016--2018. After fixing this misinterpretation we note that only a single division winner changes---in 2017 AUFNIRA should have been won by CVC4 and not by Vampire.

@@

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~\num{77352}
in~2014~\cite{CDW14} to \num{342498} in 2018.  This is partly because new
benchmarks were added to the SMT Library, but also because benchmarks
with unknown status are now eligible for the competition, and because
the competition now uses all eligible benchmarks (rather than a
randomly selected subset).

Since the re-introduction of the \ucoretrack in~2016, the
competition again has three separate tracks.  This, together with new
logics in the SMT Library, has brought the total number of competition
divisions to 115 in~2018, up from just 40 divisions
in~2014~\cite{CDW14}.  The number of job pairs increased from \num{339714}
in~2014 to~\num{1776062} in 2018, and the total wall-clock time in~2018
exceeded 7.6~years.  By these measures, the SMT Competition is one of
the largest competitions in automated reasoning.
